---
name: Shazia Ruybal Pesántez
qualifications: PhD
# position: "Postdoctoral Scientist"
# address: "London, UK"
# profilepic: "./img/headshot_2020.png"
www: shaziaruybal.com
email: "s.ruybal@imperial.ac.uk"
twitter: DrShaziaRuybal
github: shaziaruybal
orcid: "0000-0002-0495-179X"
googlescholar: "E7dY4foAAAAJ"
aboutme: "Genomic epidemiologist with expertise in population genetics, bioinformatics, and epidemiology. I'm interested in combining these approaches to population-based studies of infectious diseases, particularly malaria, and more recently COVID-19. My work has involved international collaborations in West Africa, Asia-Pacific and the Americas, and spans applied epidemiology and capacity-building in the field, to genetics and genomics in the lab, to the downstream analytics using advanced approaches and digital tools. I am an avid `R` user for data processing, visualization, communication, reports/dashboards, shiny web applications and am passionate about open science and the development of digital tools for research."
headcolor: 414141
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  vitae::awesomecv:
    page_total: true
    keep_tex: true
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(vitae)
library(kableExtra)
```

```{r functions}
readExcelSheets <- function(filename){
  # read all sheets in .xlsx 
  all <- readxl::excel_sheets(filename)
  
  # import each sheet into a list using readxl::read_excel
  list <- lapply(all, function(x) readxl::read_excel(filename, sheet = x))
  
  # save sheet name for each sheet (list)
  names(list) <- all
  
  # breaks up list and creates a dataframe for each sheet with df names matching sheet names
  list2env(list, envir = .GlobalEnv)
}
```

```{r load data, include=FALSE}
# appointments <- read_csv(here("data", "appointments.csv"))
# education <- read_csv(here("data", "education.csv"))
# research <- read_csv(here("data", "research_experience.csv"))
# teaching <- read_csv(here("data", "teaching_supervision.csv"))
# awards <- read_csv(here("data", "awards.csv"))
# leadership <- read_csv(here("data", "leadership.csv"))
# presentations <- read_csv(here("data", "presentations.csv"))
# training <- read_csv(here("data", "training.csv"))
# skills <- read_csv(here("data", "skills_table.csv"))
# tools <- read_csv(here("data", "digital_tools.csv"))

readExcelSheets(here::here("data", "cv_data.xlsx"))
```

```{r eval=FALSE, include=FALSE}
# To update CV on my website, run the following code in the console:
# rmarkdown::render("cv-vitae/cv-vitae.Rmd", output_file = "~/Documents/SR/website-apero/static/files/cv-vitae.pdf")
```

# Current Appointments
```{r}
appointments %>% 
  filter(include == "yes") %>% 
  detailed_entries(what = role,
                   when = glue::glue("{start_date}--{end_date}"),
                   with = institution,
                   where = where,
                   why = details,
                   .protect = FALSE)
```

# Education
```{r}
education %>% 
  detailed_entries(what = degree, 
                   when = year, 
                   with = institution, 
                   where = where, 
                   why = details,
                   .protect = FALSE)
```

Research Experience
====================================================
```{r}
research %>%
  detailed_entries(what = role,
                   when = dates, #glue::glue("{start_date}--{end_date}"),
                   with = institution,
                   where = where,
                   why = details,
                   .protect = FALSE)
```

<!-- \newpage -->
# Teaching and Supervision
\footnotesize
In addition to the formal supervision roles listed below, I am also actively involved in the supervision, mentorship and capacity-building of researcher officers and technical staff in partner institutions in Indonesia, Papua New Guinea and Ecuador. I am also serving on 2 PhD committees at the University of Melbourne. 

\normalsize
```{r}
teaching %>%
  detailed_entries(what = institution,
                   when = dates, #glue::glue("{start_date}--{end_date}"),
                   with = role,
                   where = where,
                   why = details,
                   .protect = FALSE)
```

# Leadership, Service & Community Engagement
\footnotesize
*Manuscript peer-review*: Nature Communications (1), Journal of Infectious Diseases (1), PLOS Genetics (1), Malaria Journal (2), International Journal for Parasitology (1), Frontiers in Genetics (2), Infection, Genetics and Evolution (1), Scientific Reports (2), Annals of Global Health (1)
\normalsize
```{r}
leadership %>%
  detailed_entries(what = glue::glue("\\textbf{<role>}: <institution> (<where>)",
                                     .open = "<", .close = ">"),
                   why = details,
                   when = dates, #glue::glue("{start_date}--{end_date}"), 
                   .protect = FALSE)

# leadership <- leadership %>% mutate(details = replace_na(details, " "))

# leadership %>%
#   brief_entries(what = glue::glue("\\textbf{<role>} \\textit{(<institution>)}. <details>", .open = "<", .close = ">"),
#                 when = dates, .protect = FALSE)
```

\newpage
# Additional training and professional development
```{r}
training %>%
  brief_entries(what = glue::glue("\\textbf{<training>} (<institution>)", .open = "<", .close = ">"),
                when = year, 
                .protect = FALSE)
```

# Skills
```{r}
# skills <- read_csv(here("data", "skills.csv"))
# skills %>%
#   detailed_entries(with = type,
#                    why = skill)
# 
skills %>%
  kable(booktabs = TRUE, format = "latex") %>%
  kable_styling(full_width = F, font_size = 8, latex_options = "hold_position") %>%
  row_spec(0, bold = T)
```

<!-- \newpage -->
# Awards and Funding
```{r}
# awards %>% 
#   detailed_entries(what = institution,
#                    when = year,
#                    with = award,
#                    why = details)

# need to manually slice awards df for now to avoid massive page break
awards %>%
  slice(1:11) %>% 
  brief_entries(glue::glue("\\textbf{<award>} (<institution>): <details>", .open = "<", .close = ">"),
                paste0(year, "\\newline~\\newline"), .protect = FALSE)
```

\newpage
```{r}
awards %>%
  slice(12:n()) %>% 
  brief_entries(glue::glue("\\textbf{<award>} (<institution>): <details>", .open = "<", .close = ">"),
                paste0(year, "\\newline~\\newline"), .protect = FALSE)
```

# Publications
\footnotesize
```{r}
# did not like this format so will add manually for now
# bibliography_entries(here("data", "papers.bib")) 
```

*indicates equal contribution

\setlength{\leftskip}{0cm}
**2023**

\setlength{\leftskip}{1cm}
Wei, H., Angrisano, F., Eriksson, E., Mazhari, R., Van, H., Zheng, S., Center, R., Boo, I., McMahon, J., Lau, J., Kiernan-Walker, N., **Ruybal-Pesántez, S.,** Mueller, I., Robinson, L. J., Anderson, D., Drummer, H. (2023). *Serological assays to measure dimeric IgA antibodies in SARS-CoV-2 infections.* Immunology & Cell Biology. http://doi.org/10.1111/imcb.12682 

Tiedje, K. E., Zhan, Q., **Ruybal‐Pesántez, S.**, Tonkin-Hill, G., He, Q., Tan, M.H., Argyropoulos, D. C., Deed, S. L., Ghansah, A., Bangre, O., Oduro, A.R., Koram, K. A., Pascual, M. & Day, K. P.(2023, *pre-print*). *Measuring changes in Plasmodium falciparum census population size in response to sequential malaria control interventions.* medRxiv. https://doi.org/10.1101/2023.05.18.23290210

Eriksson, E.M., Hart, A., Forde, M., Foroughi, S., Kiernan-Walker, N., MAzhari, R., Lucas, E.C., Margetts, M., Farchione, A., Sheerin, D., Ashdown, G., Evans, R., Chen, C., **Ruybal‐Pesántez, S.**, Conway, E., Barrios, M.H., Cornish, J., Edmonds, M., Henneken, L.M., Ioannidis, L.J., Olechnowicz, S.W.Z., Munnings, R.B., Groom, J. R., Hansen, D.S., Bowden, R., Coussens, A.K., Tye-Din, J.A., Bryan, V.L., Mueller, I. (2023, *pre-print*). *Cohort Profile: A longitudinal Victorian COVID-19 cohort (COVID PROFILE).* medRxiv. https://doi.org/10.1101/2023.04.27.23289157

**Ruybal‐Pesántez, S.**, Sáenz, F.S., Deed, S. L., Johnson, E. K., Larremore, D. B., Vera-Arias, C. A., Tiedje, K. E., Day, K. P. (2023). *Molecular epidemiology of continued Plasmodium falciparum disease transmission after an outbreak in Ecuador.* Frontiers in Tropical Diseases. https://doi.org/10.3389/fitd.2023.1085862

Mazhari, R., Takashima, E., Longley, R.J., **Ruybal‐Pesántez, S.**, White, M. T., Kanoi, B. N., Nagaoka, H., Kiniboro, B., Siba, P., Tsuboi, T., Mueller, I. (2023). *Identification of novel Plasmodium vivax proteins associated with protection against clinical malaria.* Frontiers in Cellular and Infection Microbiology (Biology and Pathogenicity of Non-falciparum Malaria - Vol II Research Topic). https://doi.org/10.3389/fcimb.2023.1076150

\setlength{\leftskip}{0cm}
**2022**

\setlength{\leftskip}{1cm}
Tiedje, K. E., Oduro, A. R., Bangre O., Amenga‐Etego, L., Dadzie, S. K., Appawu, M. A., Frempong, K., Asoala, V., **Ruybal‐Pesántez, S.**, Narh, C. A., Deed, S. L., Argyropoulos, D. C., Ghansah, A., Agyei, S. A., Segbaya, S., Desewu, K., Williams, I., Simpson, J. A., Malm, K., Pascual, M., Koram, K. A.,  & Day, K. P. (2022). *Indoor residual spraying with a non-pyrethroid insecticide reduces the reservoir of Plasmodium falciparum in a high-transmission area in northern Ghana.* PLOS Global Public Health. https://doi.org/10.1371/journal.pgph.0000285 

Charnaud, S., Munro, J., Semenec, L., Mazhari, R., Brewster, J., Bourke, C., **Ruybal‐Pesántez, S.**, James, R., Lautu‐Gumal, D., Karunajeewa, H., & Mueller, I. (2022). *PacBio long‐read amplicon sequencing for scalable high‐resolution population allele typing of the complex CYP2D6 locus*. Communications Biology. https://doi.org/10.1038/s42003-022-03102-8

**Ruybal‐Pesántez, S.**, Tiedje, K. E., Pilosof, S., Tonkin‐Hill, G., He, Q., Rask, T. S., Amenga‐Etego, L., Oduro, A. R., Koram, K. A., Pascual, M., & Day, K. P. (2022). *Age‐specific patterns of DBLα var diversity can explain why residents of high malaria transmission areas remain susceptible to Plasmodium falciparum blood stage infection throughout life.* International Journal for Parasitology. https://doi.org/10.1016/j.ijpara.2021.12.001

\setlength{\leftskip}{2cm}
  **- This work was featured on the Herminthology #WomenBehindTheWork initiative \href{https://facebook.com/102760015458811/posts/151514423916703/?d=n}{\faExternalLink}**

\setlength{\leftskip}{1cm}
Feng, Q., Tiedje, K. E., **Ruybal‐Pesántez, S.**, Tonkin‐Hill, G., Duffy, M. F., Day, K. P., Shim, H., & Chan, Y. (2022). *An accurate method for identifying recent recombinants from unaligned sequences.* Bioinformatics. https://doi.org/10.1093/bioinformatics/btac012

\setlength{\leftskip}{0cm}
**2021** 

\setlength{\leftskip}{1cm}
**Ruybal‐Pesántez, S.**, Sáenz, F., Deed, S. L., Johnson, E. K., Larremore, D. B., Vera-Arias, C. A., Tiedje, K. E. & Day, K. P. (2021, *pre-print*). *Clinical malaria incidence following an outbreak in Ecuador was predominantly associated with Plasmodium falciparum with recombinant variant antigen gene repertoires.* medRxiv. https://doi.org/10.1101/2021.04.12.21255093

Mazhari, R., **Ruybal‐Pesántez, S.**, Angrisano, F., Kiernan‐Walker, N., Hyslop, S., Longley, R. J., Bourke, C., Chen, C., Williamson, D. A., Robinson, L. J., Mueller, I., & Eriksson, E. M. (2021). *SARS‐CoV‐2 Multi‐Antigen Serology Assay*. Methods and Protocols, 4(4), 72. https://doi.org/10.3390/mps4040072

Argyropoulos, D. C.* , **Ruybal‐Pesántez, S.*** , Deed, S. L., Oduro, A. R., Dadzie, S. K., Appawu, M. A., Asoala, V., Pascual, M., Koram, K. A., Day, K. P., & Tiedje, K. E. (2021). *The impact of indoor residual spraying on Plasmodium falciparum microsatellite variation in an area of high seasonal malaria transmission in Ghana, West Africa*. Molecular Ecology, 30(16), 3974–3992. https://doi.org/10.1111/mec.16029

\setlength{\leftskip}{2cm}
  **- This work was chosen by the editors to be featured in the Molecular Ecology blog \href{https://molecularecologyblog.com/2021/09/01/interview-with-the-authors-does-indoor-spraying-alter-the-genetic-diversity-of-malaria-causing-parasites-and-what-does-this-mean-for-long-term-control/}{\faExternalLink}**

\setlength{\leftskip}{1cm}
Tonkin‐Hill, G., **Ruybal‐Pesántez, S.**, Tiedje, K. E., Rougeron, V., Duffy, M. F., Zakeri, S., Pumpaibool, T., Harnyuttanakorn, P., Branch, O. H., Ruiz‐Mesía, L., Rask, T. S., Prugnolle, F., Papenfuss, A. T., Chan, Y., & Day, K. P. (2021). *Evolutionary analyses of the major variant surface antigen‐encoding genes reveal population structure of Plasmodium falciparum within and between continents*. PLOS Genetics, 17(2), e1009269. https://doi.org/10.1371/journal.pgen.1009269

\setlength{\leftskip}{2cm}
  **- This work was chosen by the editors to be featured with an accompanying Perspectives piece \href{https://doi.org/10.1371/journal.pgen.1009344}{\faExternalLink}**


\setlength{\leftskip}{0cm}
**2020**

\setlength{\leftskip}{1cm}
Narh, C. A., Ghansah, A., Duffy, M. F., **Ruybal‐Pesántez, S.**, Onwona, C. O., Oduro, A. R., Koram, K. A., Day, K. P.* , & Tiedje, K. E.* (2020). *Evolution of antimalarial drug resistance markers in the reservoir of Plasmodium falciparum infections in the Upper East Region of Ghana*. The Journal of Infectious Diseases. https://doi.org/10.1093/infdis/jiaa286

\setlength{\leftskip}{0cm}
**2019**

\setlength{\leftskip}{1cm}
Pilosof, S., He, Q., Tiedje, K. E., **Ruybal‐Pesántez, S.**, Day, K. P., & Pascual, M. (2019). *Competition for hosts modulates vast antigenic diversity to generate persistent strain structure in Plasmodium falciparum.* PLOS Biology, 17(6), e3000336. https://doi.org/10.1371/journal.pbio.3000336

\setlength{\leftskip}{0cm}
**2018**

\setlength{\leftskip}{1cm}
He, Q., Pilosof, S., Tiedje, K. E., **Ruybal‐Pesántez, S.**, Artzy‐Randrup, Y., Baskerville, E. B., Day, K. P., & Pascual, M. (2018). *Networks of genetic similarity reveal non‐neutral processes shape strain structure in Plasmodium falciparum*. Nature Communications, 9(1), 1817. https://doi.org/10.1038/s41467-018-04219-3

Rorick, M. M., Artzy‐Randrup, Y., **Ruybal‐Pesántez, S.**, Tiedje, K. E., Rask, T. S., Oduro, A., Ghansah, A., Koram, K., Day, K. P., & Pascual, M. (2018). *Signatures of competition and strain structure within the major blood‐stage antigen of Plasmodium falciparum in a local community in Ghana*. Ecology and Evolution, 8(7), 3574–3588. https://doi.org/10.1002/ece3.3803

\setlength{\leftskip}{0cm}
**2017**

\setlength{\leftskip}{1cm}
**Ruybal‐Pesántez, S.**, Tiedje, K. E., Rorick, M. M., Amenga‐Etego, L., Ghansah, A., Oduro, A. R., Koram, K. A., & Day, K. P. (2017). *Lack of Geospatial Population Structure Yet Significant Linkage Disequilibrium in the Reservoir of Plasmodium falciparum in Bongo District, Ghana*. The American Journal of Tropical Medicine and Hygiene, 97(4), 1180–1189. https://doi.org/10.4269/ajtmh.17-0119

**Ruybal‐Pesántez, S.*** , Tiedje, K. E.* , Tonkin‐Hill, G., Rask, T. S., Kamya, M. R., Greenhouse, B., Dorsey, G., Duffy, M. F., & Day, K. P. (2017). *Population genomics of virulence genes of Plasmodium falciparum in clinical isolates from Uganda*. Scientific Reports, 7(1), 11810. https://doi.org/10.1038/s41598-017-11814-9

\setlength{\leftskip}{0cm}

<!-- \newpage -->
# Digital tools
\footnotesize
For other non-traditional academic contributions, I have also developed several `R Shiny` web applications to support COVID-19 surveillance efforts and `R flexdashboard` for real-time updates and data visualization of both programmatic/operational aspects and preliminary epidemiological trends as part of the coordination of population-based field studies. [Check out my GitHub \faExternalLink](https://github.com/shaziaruybal)
```{r}
tools %>%
  detailed_entries(what = authors,
                   when = year,
                   with = tool,
                   why = details,
                   .protect = FALSE)

# tools %>% 
#   brief_entries(glue::glue("\\textbf{<tool>}, (Authors: <authors>): <details>", .open = "<", .close = ">"),
#                 paste0(year, "\\newline~\\newline"), .protect = FALSE)
```

<!-- \newpage -->
# Selected presentations
\footnotesize
I have participated in oral and poster presentations at >20 conferences (15 international, 5 national; 6 travel awards). 
```{r}
presentations <- presentations %>% filter(include == "yes")

presentations %>%
  brief_entries(glue::glue("\\textbf{<topic>}, <where>", .open = "<", .close = ">"),
                paste0(year, "\\newline~\\newline"), 
                .protect = FALSE)
```
